\documentclass[a4paper,10pt,fleqn]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amssymb}
\usepackage{adjustbox}
\usepackage{natbib}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{fuzz}
\usepackage{schemapk}
\usepackage{zed-maths}
\usepackage{zed-proof}
\newdimen\savedleftskip
\begin{document}

\section*{Basic Generic Definitions}
\addcontentsline{toc}{section}{Basic Generic Definitions}

\subsection*{Example 1 : Generic Identity Function}
\addcontentsline{toc}{subsection}{Example 1 : Generic Identity Function}

\noindent The identityGB function works for any type:

\bigskip

\begin{gendef}[X]
  identityGB: X \fun X
\where
  \forall x : X @ identityGB(x) = x
\end{gendef}

\noindent This defines identityGB polymorphically. It can be used with any type: identityGB applied to a natural number returns that number, identityGB applied to a set returns that set, etc.

\bigskip

\subsection*{Example 2 : Generic Constant}
\addcontentsline{toc}{subsection}{Example 2 : Generic Constant}

\noindent A generic empty set:

\bigskip

\begin{gendef}[X]
  empty: \power X
\where
  empty = \{\}
\end{gendef}

\noindent The empty set can be instantiated for any type X.

\bigskip

\subsection*{Example 3 : Generic Pair Functions}
\addcontentsline{toc}{subsection}{Example 3 : Generic Pair Functions}

\noindent Extract the first element of a pair:

\bigskip

\begin{gendef}[X, Y]
  fst: X \cross Y \fun X
\where
  \forall x : X; y : Y @ fst(x, y) = x
\end{gendef}

\noindent Extract the second element:

\bigskip

\begin{gendef}[X, Y]
  snd: X \cross Y \fun Y
\where
  \forall x : X; y : Y @ snd(x, y) = y
\end{gendef}

\noindent These work for pairs of any types.

\bigskip

\subsection*{Example 4 : Generic Singleton Set}
\addcontentsline{toc}{subsection}{Example 4 : Generic Singleton Set}

\noindent Create a set containing a single element:

\bigskip

\begin{gendef}[X]
  singleton: X \fun \power X
\where
  \forall x : X @ singleton(x) = \{x\}
\end{gendef}

\noindent Given any value, singleton produces a set containing just that value.

\bigskip

\subsection*{Example 5 : Generic Sequence Operations}
\addcontentsline{toc}{subsection}{Example 5 : Generic Sequence Operations}

\noindent The head of a sequence (generic version):

\bigskip

\begin{gendef}[X]
  headOf: \seq~X \fun X
\where
  \forall s : \seq~X | \# s > 0 @ headOf(s) = s(1)
\end{gendef}

\noindent This is defined only for non-empty sequences.

\bigskip

\subsection*{Example 6 : Generic Set Membership ( Conceptual )}
\addcontentsline{toc}{subsection}{Example 6 : Generic Set Membership ( Conceptual )}

\noindent Set membership can be tested using the built-in 'in' operator in Z notation. In other contexts, you might define a membership test function, but Z notation provides this as a primitive predicate: x in S is true exactly when x is an element of S.

\bigskip

\subsection*{Example 7 : Generic Optional Type ( Conceptual )}
\addcontentsline{toc}{subsection}{Example 7 : Generic Optional Type ( Conceptual )}

\noindent Optional values can be modeled with free types in Z notation. For a specific type, you would define: Option ::= none $\mid$ some⟨T⟩. Generic optional types require more complex setup beyond basic gendef blocks, as free types in Z are typically defined for specific types rather than being polymorphic.

\bigskip

\subsection*{Example 8 : Generic Swap Function}
\addcontentsline{toc}{subsection}{Example 8 : Generic Swap Function}

\noindent Swap the components of a pair:

\bigskip

\begin{gendef}[X, Y]
  swap: X \cross Y \fun Y \cross X
\where
  \forall x : X; y : Y @ swap(x, y) = (y, x)
\end{gendef}

\noindent Converts (x, y) to (y, x).

\bigskip

\subsection*{Example 9 : Generic Cardinality}
\addcontentsline{toc}{subsection}{Example 9 : Generic Cardinality}

\noindent Size of a set (for finite sets):

\bigskip

\begin{gendef}[X]
  size: \power X \fun \nat
\where
  \forall S : \power X @ size(S) = \# S
\end{gendef}

\noindent Returns the number of elements in a finite set.

\bigskip

\subsection*{Example 10 : Generic List Construction}
\addcontentsline{toc}{subsection}{Example 10 : Generic List Construction}

\noindent Construct a singleton sequence:

\bigskip

\begin{gendef}[X]
  single: X \fun \seq~X
\where
  \forall x : X @ single(x) = \langle x \rangle
\end{gendef}

\noindent Wraps a value in a single-element sequence.

\bigskip

\subsection*{Example 11 : Using Generic Definitions}
\addcontentsline{toc}{subsection}{Example 11 : Using Generic Definitions}

\noindent Once defined, generic functions can be instantiated for specific types:

\bigskip

\begin{axdef}
id\_nat : \nat \fun \nat \\
id\_set : \power \nat \fun \power \nat
\where
id\_nat = identityGB[\nat] \\
id\_set = identityGB[\power \nat]
\end{axdef}

\noindent Here identityGB[N] is the identityGB function instantiated for natural numbers, and identityGB[P N] is instantiated for sets of natural numbers.

\bigskip

\subsection*{Example 12 : Generic Constants}
\addcontentsline{toc}{subsection}{Example 12 : Generic Constants}

\noindent Define a generic empty sequence:

\bigskip

\begin{gendef}[X]
  emptySeq: \seq~X
\where
  emptySeq = \langle \rangle
\end{gendef}

\noindent This can be instantiated as emptySeq[N] for an empty sequence of naturals, emptySeq[Person] for an empty sequence of persons, etc.

\bigskip

\end{document}