\documentclass[a4paper,10pt,fleqn]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amssymb}
\usepackage{adjustbox}
\usepackage{natbib}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{fuzz}
\usepackage{zed-maths}
\usepackage{zed-proof}
\newdimen\savedleftskip
\title{Foreign Keys in Relational Schemas}
\author{txt2tex example}
\hypersetup{pdftitle={Foreign Keys in Relational Schemas},pdfauthor={txt2tex example},pdfcreator={txt2tex v1.3.1}}
\begin{document}

\maketitle

\begin{zed}[RecipeId, IngredientId, EmployeeId, CuisineTag, IngredientName, RecipeName, EmployeeName, PositionName]\end{zed}

\medskip

\section*{Pattern 1: Single-Attribute Primary Key}

\noindent Each relation has exactly one pk line.  txt2tex underlines the attribute in the schema box and emits a PK statement immediately after.

\bigskip

\begin{schema}{Recipe}
recipeId : RecipeId \\
name : RecipeName \\
cuisine : CuisineTag
\end{schema}
\noindent$\mathrm{PK}(\mathrm{Recipe}) = \{recipeId\}$

\smallskip

\begin{schema}{Ingredient}
ingredientId : IngredientId \\
name : IngredientName
\end{schema}
\noindent$\mathrm{PK}(\mathrm{Ingredient}) = \{ingredientId\}$

\smallskip

\medskip

\section*{Pattern 2: Composite Primary Key}

\noindent Two pk lines in one schema body.  Both attributes are underlined; txt2tex collapses them into one PK set: PK(RecipeIngredient) = $\{recipeId, ingredientId\}$.

\bigskip

\begin{schema}{RecipeIngredient}
recipeId : RecipeId \\
ingredientId : IngredientId
\end{schema}
\noindent$\mathrm{PK}(\mathrm{RecipeIngredient}) = \{recipeId, ingredientId\}$

\smallskip

\medskip

\section*{Pattern 3: Single-Attribute Foreign Key}

\noindent A foreign key constraint says every value of the referencing attribute must appear in the referenced relation.  No special keyword is needed: write the constraint as a Z predicate using forall/exists or projection with subset.

\bigskip

\noindent RecipeIngredient.recipeId must reference an existing Recipe:

\bigskip

\noindent
$\forall ri : RecipeIngredient @ \exists r : Recipe @ ri.recipeId = r.recipeId$


\noindent Equivalently, using projection and subset:

\bigskip

\noindent
$\mathrm{Project}_{recipeId}(RecipeIngredient) \subseteq \mathrm{Project}_{recipeId}(Recipe)$


\noindent RecipeIngredient.ingredientId must reference an existing Ingredient:

\bigskip

\noindent
$\forall ri : RecipeIngredient @ \exists i : Ingredient @ ri.ingredientId = i.ingredientId$


\medskip

\section*{Pattern 4: Composite FK and Self-Referential FK}

\noindent Both FK constraints for RecipeIngredient can be stated together in a single zed block.  Each predicate is separated by an also command in the output.

\bigskip

\begin{zed}
\forall ri : RecipeIngredient @ \exists r : Recipe @ ri.recipeId = r.recipeId
\also
\forall ri : RecipeIngredient @ \exists i : Ingredient @ ri.ingredientId = i.ingredientId
\end{zed}

\noindent Employee carries a managerId that references another row in the same relation.  The self-referential FK constraint:

\bigskip

\begin{schema}{Employee}
employeeId : EmployeeId \\
name : EmployeeName \\
position : PositionName \\
managerId : EmployeeId
\end{schema}
\noindent$\mathrm{PK}(\mathrm{Employee}) = \{employeeId\}$

\smallskip

\noindent Every managerId in Employee must match a valid employeeId in the same relation:

\bigskip

\begin{zed}
\forall e : Employee @ \exists m : Employee @ e.managerId = m.employeeId
\end{zed}

\end{document}